---
title: "Outbreak Analyses"
author: "McCaide Wooten"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load Libraries#
library(tidyverse)
library(stats)
library(MASS)
library(ggthemes)
library(ggsci)
library(readxl)
library(dplyr)
library(readr)
library(glmulti)
library(ggplot2)
library(survival)
library(reshape2)
library(GlmSimulatoR)
library(AER)
library(kableExtra)
library(Hmisc)
library(knitr)
library(vtable)
library(foreign)
library(corrplot)
library(MuMIn)
library(glmulti)
library(table1)
library(flextable)
library(beepr)
library(aod)
library(extrafont)
library(car)        
library(ResourceSelection)
library(pROC)

```

```{r Population Structure Plotting}
### Definitive & Presumptive Cases ###

#Load Data#
Data <- read_csv("C:/INSERT FILE DIRECTORY/Carcasses.csv")

#Data Modification
AgeSexCases <- Data %>%
  group_by(Age) %>%                     
  filter(!is.na(Age), Sex != "U",`Case` != "N") %>%
  ungroup()

AgeSexCases$Age <- as.numeric(AgeSexCases$Age)

AgeSexCases_Grouped <- AgeSexCases %>%
  mutate(AgeGroup = cut(Age,
  breaks = c(0,.99,1.01,2.01,3.01,4.01,Inf), # Define age group ranges
  labels = c("<1", "1", "2", "3", "4", "5+"),
  right = FALSE
  ))

# Aggregate data by age group and sex
agg_AgeSexCases <- AgeSexCases_Grouped %>%
  count(AgeGroup, Sex) %>% # Count entries per age group and gender
  mutate(population = ifelse(Sex == "M", -n, n)) # Make males negative for the pyramid

# Create the population pyramid
pyramid_plot_cases <- ggplot(agg_AgeSexCases, aes(x = AgeGroup, y = population, fill = Sex)) +
  geom_bar(stat = "identity", width = 0.85, color = "black", linewidth = 0.2) + # Thin borders
  scale_y_continuous(labels = abs,              # Show absolute values on y-axis
                     breaks = seq(-25, 25, 5),
                     limits = c(-30, 30) # Adjust limits for symmetry)
  ) +
  coord_flip() +                                # Flip coordinates for horizontal bars
  labs(
    title = "Definitive & Presumptive Cases",
    x = "Age (Years)",
    y = "Count",
    fill = NULL # Remove legend title
  ) +
  scale_fill_manual(values = c("M" = "#2166AC", "F" = "#B2182B")
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(size = 12, face = "plain"),                            
    axis.title.x = element_text(size = 12, face = "plain"),  # Plain x-axis title
    axis.title.y = element_text(size = 12, face = "plain"),  # Plain y-axis title
    axis.text = element_text(size = 10),                    # Smaller tick labels
    legend.position = "right",                                 # Legend on right
    legend.text = element_text(size = 10),                   # Smaller legend text
    panel.grid.major = element_blank(),                      # No gridlines
    panel.grid.minor = element_blank()                       # No minor gridlines
  )

# Display the plot
print(pyramid_plot_cases)



### All Sampled Carcasses ###

#Data Modification
AgeSexCleaned <- Data %>%
  group_by(Age) %>%                     
  filter(!is.na(Age), Sex != "U", Sampled != "N") %>%
  ungroup()

AgeSexCleaned$Age <- as.numeric(AgeSexCleaned$Age)

AgeSexCleaned_Grouped <- AgeSexCleaned %>%
  mutate(AgeGroup = cut(Age,
  breaks = c(0,.99,1.01,2.01,3.01,4.01,Inf), # Define age group ranges
  labels = c("<1", "1", "2", "3", "4", "5+"),
  right = FALSE
  ))

# Aggregate data by age group and sex
agg_AgeSexCleaned <- AgeSexCleaned_Grouped %>%
  count(AgeGroup, Sex) %>% # Count entries per age group and gender
  mutate(population = ifelse(Sex == "M", -n, n)) # Make males negative for the pyramid

# Create the population pyramid
pyramid_plot_all <- ggplot(agg_AgeSexCleaned, aes(x = AgeGroup, y = population, fill = Sex)) +
  geom_bar(stat = "identity", width = 0.85, color = "black", linewidth = 0.2) + # Thin borders
  scale_y_continuous(labels = abs,              # Show absolute values on y-axis
                     breaks = seq(-100, 100, 20),
                     limits = c(-120, 120) # Adjust limits for symmetry)
  ) +
  coord_flip() +                                # Flip coordinates for horizontal bars
  labs(
    title = "All Sampled Carcasses",
    x = "Age (Years)",
    y = "Count",
    fill = NULL # Remove legend title
  ) +
  scale_fill_manual(values = c("M" = "#2166AC", "F" = "#B2182B")
  ) +
  theme_classic(base_size = 12) + 
  theme(
    plot.title = element_text(size = 12, face = "plain"),                            
    axis.title.x = element_text(size = 12, face = "plain"),  # Plain x-axis title
    axis.title.y = element_text(size = 12, face = "plain"),  # Plain y-axis title
    axis.text = element_text(size = 10),                    # Smaller tick labels
    legend.position = "right",                                 # Legend on right
    legend.text = element_text(size = 10),                   # Smaller legend text
    panel.grid.major = element_blank(),                      # No gridlines
    panel.grid.minor = element_blank()                       # No minor gridlines
  )

# Display the plot
print(pyramid_plot_all)
```


```{r Chi Square Comparisons of Live Population Observations to Calculated "Norms"}

# Load the dataset
chi_data <- read_excel("C:/INSERT FILE DIRECTORY/ChiSquare.xlsx", sheet = "Sheet1")

# Group by Analysis, Year, Observational Unit, and Location, then run chi-square tests
chi_results <- chi_data %>%
  group_by(Analysis, Location, Observational_Unit, Year) %>%
  summarise(
    Statistic = chisq.test(Observed, p = Expected / sum(Expected))$statistic,
    DF = chisq.test(Observed, p = Expected / sum(Expected))$parameter,
    P_Value = chisq.test(Observed, p = Expected / sum(Expected))$p.value,
    .groups = "drop"
  )

# Print results
print(chi_results)

# Male:Female Case Analysis
observed <- c(20, 10)
expected <- c(13, 17)

chisq_testMF <- chisq.test(x = observed, p = expected/sum(expected), rescale.p = TRUE)

print(chisq_testMF)